### 2023-02-10 article.tex
Add code availablity

\paragraph{Simple example}

% Alex and Blake?
Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that describe an experiment with computational steps.
Both manuscripts are accepted by an equally prestigous journal after peer view. 
One researcher, however, does not supply the
computer code (from now on: 'code') that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses why code, a type of paradata, 
should be supplied
and how it should be supplied,
and claims that code is just as important as the paper itself.

\paragraph{The first use of paradata}

@@

\paragraph{Code is paradata}

The code used in computational experiments is just that, 
'data about the data collection 
process', as it commonly downloads data, selects relevant subsets in those
downloaded data and performs statistical tests to generate results.
Code, hence, is paradata and this chapter explores and illustrates the
consequence of that premise.

\paragraph{Making paradata useful}

A recent paper explores the use of paradata to increase the impact of data,
stating that lack of paradata can be seen as 'a drastic constraint'
in the use of data and offer some suggestions to 
make paradata useful for data re(use):
paradata should be comprehensive, documented in a useful way, 
the documention and data should have co-evolved, 
and the paradata should be good to computers \cite{huvila2022improving}.
These suggestions appear to work well for code too
and will be discussed in detail below.

\paragraph{The goal of good paradata}

@@

@@

\paragraph{Introduction}

This paper uses genetic epidemiology as a specific example,
to illustrate that the code is paradata,
and that this type of paradata is relevant.
However, any field that uses computation 
on sensitive data in its experiments,
could be used as an example.

\paragraph{How commonly code is published}\label{para:how-commonly-code-is-published}

Going back to the example scientists in the introduction,
where one does and the other does not publish the code,
it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

\paragraph{For some fields, the experiment is actually run by code}

@@

\paragraph{The experiments within genetic epidemology works are done by code}

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
All these steps are described in the paper (\cite{ahsan2017relative}), 
but the code itself is unpublished
and the paper does not mention any way to obtain it.
One could send an email to the correspending,
yet it is known that the reponse rate of correnspondence authors 
with a request for data in related fields is around 50\% 
\cite{manca2018non, teunis2015corresponding} 
(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).

\paragraph{The goal of useful paradata is reproducible science}

@@

\paragraph{Sensitive data}

@@

\paragraph{Code holds the ground truth}

@@

\paragraph{Code should be hosted}

@@

\paragraph{Code must have extensive error handling}

@@

\paragraph{Using a tool for coding style}

@@

\paragraph{Using a tool for complexity}

@@

\paragraph{Code must be tested}

@@

\paragraph{Adding a license}

@@

\paragraph{Test-driven development}

@@

\paragraph{Code must be reviewed to ensure the correct amount of documentation}

@@

\paragraph{Publishing a virtual environment}

@@

\paragraph{Hosted code keeps a history of changes}

@@

\paragraph{Invest in training}

@@

\paragraph{Code needs maintanance}

@@

\paragraph{Authors can be contacted}

@@

\paragraph{How to handle bugs?}

@@

% \paragraph{Still need the same data to do a reproducible experiment}

% \paragraph{Recommendations}

@@

\paragraph{Final goal: runnable code}

@@

\paragraph{First step: publish code alongside papers}

>>>>

% \paragraph{Simple example}

% Alex and Blake?
Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that describe an experiment with computational steps.
Both manuscripts are accepted by an equally prestigious journal after peer review. 
One researcher, however, does not supply the
computer code (from now on: 'code') that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses why code, a type of paradata, 
should be supplied
and how it should be supplied,
and claims that code is just as important as the paper itself.

% \paragraph{The first use of paradata}

@@

% \paragraph{Code is paradata}

The code used in computational experiments is just that, 
'data about the data collection 
process', as it commonly downloads data,
selects relevant subsets in those data,
performs statistical tests and generates figures.
In each case, code answers the question: 
'Where is it (i.e. the result) coming from?'.
Code, hence, is paradata and this chapter explores and illustrates the
consequence of that premise.

% \paragraph{Making paradata useful}

A recent paper explores the use of paradata to increase the impact of data,
stating that lack of paradata can be seen as 'a drastic constraint'
in the use of data and offer some suggestions to 
make paradata useful for data re(use):
paradata should be comprehensive, documented in a useful way, 
the documention and data should have co-evolved, 
and the paradata should be computer-friendly \cite{huvila2022improving}.
At first glance, these suggestions appear to work well for code too
and will be discussed in detail below.

% \paragraph{The goal of good paradata}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code availability}\label{sec:code-availability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For arcticles in Science magazine (years 2011-2012), 24 out of 204
studies published the code \cite{stodden2018empirical}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

% \paragraph{Obtaining code by email}

When code of an academic paper is not published, 
one can contact the correspending author and request it.
However, the reponse rate of correnspondence authors 
with a request for data in computational fields is around 50\% 
\cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
When getting an answer of the corresponding author, 
half (65 out of 134) will actually share the code \cite{stodden2018empirical}.
The responses of unwilling authors (see \cite{stodden2018empirical} for 
some real examples) can come across as so caustic, 
that one is excused from not contacting an a corresponding author.

@@

% \paragraph{Introduction}

This paper uses genetic epidemiology as a specific example,
to illustrate that the code is paradata,
and that this type of paradata is relevant.
However, any field that uses computation 
and sensitive data in its experiments,
could be used as an example.

% \paragraph{For some fields, the experiment is actually run by code}

@@

% \paragraph{The experiments within genetic epidemology works are done by code}

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
All these steps are described in the paper (\cite{ahsan2017relative}), 
but the code itself is unpublished
and the paper does not mention any way to obtain it.
Sending an email is known to be unlikely to work 
(see section \ref{sec:code-availability}).

% \paragraph{The goal of useful paradata is reproducible science}

@@

% \paragraph{Sensitive data}

@@

% \paragraph{Code holds the ground truth}

@@

% \paragraph{Code should be hosted}

@@

% \paragraph{Code must have extensive error handling}

@@

% \paragraph{Using a tool for coding style}

@@

% \paragraph{Using a tool for complexity}

@@

% \paragraph{Code must be tested}

@@

% \paragraph{Adding a license}

@@

% \paragraph{Test-driven development}

@@

% \paragraph{Code must be reviewed to ensure the correct amount of documentation}

@@

% \paragraph{Publishing a virtual environment}

@@

% \paragraph{Hosted code keeps a history of changes}

@@

% \paragraph{Invest in training}

@@

% \paragraph{Code needs maintanance}

@@

% \paragraph{Authors can be contacted}

@@

% \paragraph{How to handle bugs?}

@@

% % \paragraph{Still need the same data to do a reproducible experiment}

% % \paragraph{Recommendations}

@@

% \paragraph{Final goal: runnable code}

@@

% \paragraph{First step: publish code alongside papers}
